---
title: "Income and Housing Costs by County"
author: "Parker Pratt"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cerulean
    code_folding: hide

params: 
  data:
   value: NULL
  state: 
   value: NULL
  income: 
    label: "Income"
    value: [50000, 80000]
    input: slider
    min: 25000
    max: 450000
    step: 10000
  housing: 
    label: "Housing Cost"
    value: [250000, 500000]
    input: slider
    min: 50000
    max: 2750000
    step: 25000
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(sf)
library(readxl)
library(plotly)
library(tidyverse)
```

```{r}
data = params$data
```

Below is a radar plot with all states and their average housing costs and income compared to each other.

```{r, message=FALSE}
radar_df <- data %>%
  group_by(STATE) %>%
  summarise(avg_income = mean(INCOME, na.rm = TRUE), 
            avg_housing_cost = mean(COST, na.rm = TRUE))

radar_plot <- plot_ly(
  type = 'scatterpolar',
  fill = 'toself'
) %>%
  add_trace(
    r = c(radar_df$avg_income, radar_df$avg_income[1]),
    theta = c(radar_df$STATE, radar_df$STATE[1]),
    name = "Avg Income"
  ) %>%
  add_trace(
    r = c(radar_df$avg_housing_cost, radar_df$avg_housing_cost[1]),
    theta = c(radar_df$STATE, radar_df$STATE[1]),
    name = "Avg Housing Cost"
  ) %>%
  layout(
    polar = list(
      radialaxis = list(
        visible = TRUE,
        range = c(0, max(c(radar_df$avg_income, radar_df$avg_housing_cost)))
      )
    ),
    showlegend = TRUE
  )

radar_plot

```

Below is a map of all counties in the continental United States with their associated average income.

```{r}
custom_colors <- c("grey", "blue")
breaks <- c(0, 100000)
labels <- c("< 30000", "> 100000")
threshold <- 100000

income_plot <- ggplot(data = data) +
  geom_sf(aes(fill = INCOME)) +
  scale_fill_gradientn(name = "Yearly Personal Income ($)",
                       colors = custom_colors,
                       limits = c(0, threshold),
                       breaks = breaks,
                       labels = labels,
                       na.value = "blue") +

  labs(title = "Income is Fairly Consistent",
       subtitle = "By County in the United States") +

  theme_minimal() +
  theme(plot.title = element_text(size = 18L, hjust = 0.5),
        plot.subtitle = element_text(size = 11L, hjust = 0.5),
        panel.grid = element_blank(),
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())

print(income_plot)
```

```{r echo=FALSE, include=FALSE}
# incomeI <- ggplotly(income_plot)
# incomeI
```

Below is a map with all counties that have an average income between `r min(params$income)` and `r max(params$income)` USD. All counties which are black do not fit in the range.

```{r}
custom_colors <- c("grey", "blue")
breaks <- c(min(params$income), max(params$income))
labels <- c(min(params$income), max(params$income))
threshold <- max(params$income)

income_plot_param <- ggplot(data = data) +
  geom_sf(aes(fill = INCOME)) +
  scale_fill_gradientn(name = "Yearly Personal Income ($)",
                       colors = custom_colors,
                       limits = c(min(params$income), threshold),
                       breaks = breaks,
                       labels = labels,
                       na.value = "black") +

  labs(title = paste("Income between", min(params$income), "and", max(params$income)),
       subtitle = "On Average by County") +

  theme_minimal() +
  theme(plot.title = element_text(size = 18L, hjust = 0.5),
        plot.subtitle = element_text(size = 11L, hjust = 0.5),
        panel.grid = element_blank(),
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())

print(income_plot_param)
```

```{r echo=FALSE, include=FALSE}
#incomeI_param <- ggplotly(income_plot_param)
#incomeI_param
```

Below is a map of all counties in the continental United States with their associated average housing cost.

```{r}
custom_colors <- c("grey", "blue")
breaks <- c(0, 700000)
labels <- c("< 100000", "> 700000")
threshold <- 700000

house_plot <- ggplot(data = data) +
  geom_sf(aes(fill = COST)) +
  scale_fill_gradientn(name = "Average House Cost ($)",
                       colors = custom_colors,
                       limits = c(0, threshold),
                       breaks = breaks,
                       labels = labels,
                       na.value = "blue") +

  labs(title = "Western US is More Expensive for Housing",
       subtitle = "On Average by County") +

  theme_minimal() +
  theme(plot.title = element_text(size = 18L, hjust = 0.5),
        plot.subtitle = element_text(size = 11L, hjust = 0.5),
        panel.grid = element_blank(),
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())

print(house_plot)
```

```{r echo=FALSE, include=FALSE}
#houseI <- ggplotly(house_plot)
#houseI
```

Below is a map with all counties that have an average housing cost between `r min(params$housing)` and `r max(params$housing)` USD. All counties which are black do not fit in the range.

```{r}
custom_colors <- c("grey", "blue")
breaks <- c(min(params$housing) , max(params$housing) )
labels <- c(min(params$housing) , max(params$housing) )
threshold <- max(params$housing)

house_plot_param <- ggplot(data = data) +
  geom_sf(aes(fill = COST)) +
  scale_fill_gradientn(name = "Average House Cost ($)",
                       colors = custom_colors,
                       limits = c(min(params$housing) , threshold),
                       breaks = breaks,
                       labels = labels,
                       na.value = "black") +

  labs(title = paste("Housing Prices between", min(params$housing), "and", max(params$housing)),
       subtitle = "On Average by County") +

  theme_minimal() +
  theme(plot.title = element_text(size = 18L, hjust = 0.5),
        plot.subtitle = element_text(size = 11L, hjust = 0.5),
        panel.grid = element_blank(),
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())

print(house_plot_param)
```

```{r echo=FALSE, include=FALSE}
#houseI_param <- ggplotly(house_plot_param)
#houseI_param
```

Below is a map of all counties in the continental United States with their associated Housing/Income ratio.

```{r}
custom_colors <- c("grey", "blue")
breaks <- c(0, 10)
labels <- c("< 1", "> 10")
threshold <- 10

diff_plot <- ggplot(data = data) +
  geom_sf(aes(fill = COST/INCOME)) +
  scale_fill_gradientn(name = "Price : Income Ratio",
                       colors = custom_colors,
                       limits = c(0, threshold),
                       breaks = breaks,
                       labels = labels,
                       na.value = "blue") +

  labs(title = "The Midwest is the Best Place to Buy a House",
       subtitle = "Based on Percent of Income Spent") +

  theme_minimal() +
  theme(plot.title = element_text(size = 18L, hjust = 0.5),
        plot.subtitle = element_text(size = 11L, hjust = 0.5),
        panel.grid = element_blank(),
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())

print(diff_plot)
```

```{r echo=FALSE, include=FALSE}
#diffI <- ggplotly(diff_plot)
#diffI
```

Below is a map with all counties that have a Housing/Income ratio between `r round(min(params$housing)/min(params$income), 2)` and `r round(max(params$housing)/max(params$income), 2)`. All counties which are black do not fit in the range.

```{r}
min_ratio = min(params$housing)/min(params$income)
max_ratio = max(params$housing)/max(params$income)
min_ratio <- round(min_ratio, 2)
max_ratio <- round(max_ratio, 2)

custom_colors <- c("grey", "blue")
breaks <- c(min_ratio, max_ratio)
labels <- c(min_ratio, max_ratio)
threshold <- max_ratio

diff_plot_param <- ggplot(data = data) +
  geom_sf(aes(fill = COST/INCOME)) +
  scale_fill_gradientn(name = "Price : Income Ratio",
                       colors = custom_colors,
                       limits = c(min_ratio, threshold),
                       breaks = breaks,
                       labels = labels,
                       na.value = "black") +

  labs(title = paste("Housing/Income Ratio between", min_ratio, "and", max_ratio),
       subtitle = "Based on Average Housing Cost and Income") +

  theme_minimal() +
  theme(plot.title = element_text(size = 18L, hjust = 0.5),
        plot.subtitle = element_text(size = 11L, hjust = 0.5),
        panel.grid = element_blank(),
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())

print(diff_plot_param)
```

```{r echo=FALSE, include=FALSE}
#diffI_param <- ggplotly(diff_plot_param)
#diffI_param
```

This is a map with all counties in the state of `r params$state` that have an average housing cost between `r min(params$housing)` and `r max(params$housing)` USD and an average income between `r min(params$income)` and `r max(params$income)` USD. All counties which are black do not fit in the range.

```{r warning=FALSE}
plot_data <- data

               if (params$state != "All") {
                       plot_data <- plot_data %>% filter(STATE == params$state)
                }

                custom_colors <- c("grey", "blue")
                breaks <- c(min(params$income), max(params$income))
                labels <- c(min(params$income), max(params$income))
                threshold <- max(params$income)
                         
                plot_param <- ggplot(data = plot_data) +
                geom_sf(aes(fill = ifelse(INCOME >= params$income[1] & INCOME <= params$income[2] &
                                          COST >= params$housing[1] & COST <= params$housing[2], INCOME, NA),
                                          text = paste("County: ", NAME, "<br>",
                                                       "Income: $", INCOME, "<br>",
                                                       "Housing Cost: $", COST)))+
                     
                scale_fill_gradientn(name = "Yearly Personal Income ($)",
                                     colors = custom_colors,
                                     limits = c(min(params$income), threshold),
                                     breaks = breaks,
                                     labels = labels,
                                     na.value = "#000000") +
                theme_minimal() +
                theme(plot.title = element_text(size = 18L, hjust = 0.5),
                      plot.subtitle = element_text(size = 11L, hjust = 0.5),
                      panel.grid = element_blank(),
                      axis.line = element_blank(),
                      axis.text = element_blank(),
                      axis.ticks = element_blank())
                         
ggplotly(plot_param, tooltip = "text") %>%
     layout(showlegend = FALSE)
```

This is a table with the specific data for all of the counties in the above map.

```{r}
data_df <- as.data.frame(data)
               
               if (params$state != "All") {
                    data_df <- data_df %>% filter(STATE == params$state)
               }
               
               # Filtering based on input ranges
               data_df <- data_df %>%
                    filter(INCOME >= params$income[1] & INCOME <= params$income[2]) %>%
                    filter(COST >= params$housing[1] & COST <= params$housing[2])
               
               # Renaming and selecting columns
               data_df <- as.data.frame(data_df)
                                        data_df <- data_df %>%
                                        rename(HOUSING_COST = COST) %>%
                                        rename(COUNTY_NAME = NAME) %>%
                                             drop_na() %>%
                                        select(COUNTY_NAME, STATE, INCOME, HOUSING_COST)
               
               # Render the datatable
datatable(data_df, options = list(pageLength = 10, autoWidth = FALSE))
```

